\documentclass{article}

% Packages
\usepackage{amsfonts, amsthm, amsmath, amssymb}
\usepackage{graphicx, svg, subfigure}
\usepackage{geometry, enumitem, parskip, listings}
% \usepackage{cancel}

% Algorithm environments
\usepackage{float}  % Required by algorithm
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}

\usepackage{appendix}
\usepackage{url}
\usepackage[hidelinks=true]{hyperref}

% Document
\begin{document}

\include{style}
\title{\emph{iZprime} Framework: Pseudocode Documentation}
\author{\small{\url{https://github.com/Zprime137/iZprime}}}
\maketitle

\section*{Introduction}

The \emph{iZprime} framework is a high-performance C toolkit designed for prime number generation and algorithmic research. It features efficient implementations of classical and novel prime sieve algorithms, optimized for both educational use and computational tasks requiring large-scale prime enumeration.

This document provides formal pseudocode representations of the key algorithms implemented in the framework, accompanied by detailed explanations and complexity analyses. The pseudocode abstracts away low-level implementation details to emphasize the core mathematical logic and algorithmic structure of each method, making it accessible for theoretical study and practical implementation alike.

\tableofcontents
\newpage

\section{Basic Data Structures}

The prime sieve algorithms presented in this document rely on two fundamental data structures: an integer array (\ttt{INT\_ARRAY}) for maintaining the ordered list of discovered primes, and a bitmap (\ttt{BITMAP}) for efficiently representing candidate numbers and tracking their primality status during the sieving process.

\subsection{Integer Array (\ttt{INT\_ARRAY})}

For storing and managing the list of prime numbers generated by sieve algorithms, we define a dynamic integer array structure called \ttt{INT\_ARRAY}. This structure maintains an expandable array of integers along with metadata tracking the current count and total capacity.

A representative C definition of the \ttt{INT\_ARRAY} structure is:
\begin{verbatim}
typedef struct
{
    size_t capacity;       ///< Total capacity (max count of elements) of the array
    int count;             ///< Number of elements currently stored
    int *array;            ///< Pointer to the dynamically allocated int array
} INT_ARRAY;
\end{verbatim}
In this document, we use the following operations on \ttt{INT\_ARRAY}:
\begin{itemize}[noitemsep]
    \item \ttt{int\_array\_init()}: Initializes and returns an \ttt{INT\_ARRAY} object with a dynamically allocated array, with both \emph{capacity} and \emph{count} initialized to 0.
    \item \ttt{int\_array\_push($int\_array, \ value$)}: Appends the integer \emph{value} to the end of \emph{int\_array}, expanding capacity as needed.
\end{itemize}

\subsection{Bitmap (\ttt{BITMAP})}

For memory-efficient representation of candidate number ranges in sieve algorithms, we define a bitmap structure called \ttt{BITMAP}. This structure implements a compact bit array where each \emph{bit-index} corresponds to a candidate number, and each \emph{bit-state} (0 or 1) encodes that candidate's primality status.

A representative C definition of the \ttt{BITMAP} structure is:
\begin{verbatim}
typedef struct
{
    size_t size;        ///< Total number of bits in the bitmap
    uint8_t *data;      ///< A dynamically allocated array of bytes ((size + 7) / 8)
} BITMAP;
\end{verbatim}
In this document, we use the following operations on \ttt{BITMAP}:
\begin{itemize}[noitemsep]
    \item \ttt{bitmap\_init($size, \ initial\_value$)}: Initializes and returns a bitmap of given \emph{size} with all bits set to \emph{initial\_value}.
    \item \ttt{bitmap\_set\_all($bitmap$)}: Sets all bits in the \emph{bitmap} to 1.
    \item \ttt{bitmap\_get\_bit($bitmap, \ index$)}: Returns the bit-state at \emph{index} in the \emph{bitmap}.
    \item \ttt{bitmap\_flip\_bit($bitmap, \ index$)}: Flips the bit-state at \emph{index} in the \emph{bitmap}.
    \item \ttt{bitmap\_clear\_bit($bitmap, \ index$)}: Clears (sets to 0) the bit at \emph{index} in the \emph{bitmap}.
    \item \ttt{bitmap\_clear\_steps($bitmap, \ step, \ start, \ limit$)}: Clears bits in the \emph{bitmap} starting from \emph{start} and incrementing by \emph{step} until \emph{limit} is reached.
    \item \ttt{bitmap\_clone($bitmap$)}: Creates and returns a deep copy of the given \emph{bitmap}.
\end{itemize}

\section{Prime Sieve Algorithms}

This section presents detailed pseudocode for various prime sieve algorithms, accompanied by comprehensive complexity analyses. All algorithms generate an ordered list of all prime numbers up to a given limit $N$, though they employ different strategies and optimizations to achieve this goal.

\subsection{Sieve of Eratosthenes (SoE)} \label{alg:soe}

The Sieve of Eratosthenes, attributed to the ancient Greek mathematician Eratosthenes of Cyrene (c. 276--194 BCE), is one of the most elegant and fundamental algorithms for finding all prime numbers up to a specified limit $N$. 

The algorithm operates by iteratively marking the multiples of each prime $p \le \sqrt{N}$ as composite. Numbers that remain unmarked after all iterations are prime. This straightforward approach has been a cornerstone of computational number theory for centuries.

\begin{algorithm}[H]
    \caption{Sieve of Eratosthenes (SoE)}
    \begin{algorithmic}[1]        
        \Procedure{SoE}{$\ttt{int}: n$}{ $\rightarrow \ttt{INT\_ARRAY}$}
            \State \Comment{\textbf{1. Initialization:}}
            \State \ttt{INT\_ARRAY:} $primes \gets \Call{int\_array\_init}$ \algocomment{initialize primes array}
            \State \Call{int\_array\_push}{$primes, 2$} \algocomment{append 2 to primes, to focus on odd candidates}

            \State \ttt{BITMAP:} $sieve \gets \Call{bitmap\_init}{n + 1, 1}$ \algocomment{create sieve bitmap with all bits set to 1}

            \State \Comment{\textbf{2. Sieve Process: mark composites and collect primes}}
            \State $n\_sqrt \gets \lceil\sqrt{n}\rceil$ \algocomment{compute root limit}
            \For{$i \gets 3$ to $n$ with step 2} \algocomment{iterate over odd candidates}
                \If{\Call{bitmap\_get\_bit}{$sieve, i$}} \algocomment{if $i$ is prime}
                    \State \Call{int\_array\_push}{$primes, i$} \algocomment{append $i$ to primes}
                    \If{$i < n\_sqrt$} \algocomment{if $i$ is root prime}
                        \State \Call{bitmap\_clear\_steps}{$sieve, 2i, i^2, n$} \algocomment{mark its odd multiples starting from $i^2$}
                    \EndIf
                \EndIf
            \EndFor

            \State \Comment{\textbf{3. Return $primes$}}
            \State \Return \emph{primes}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\subsubsection*{Complexity Analysis}

\paragraph{Space Complexity:} Since SoE maintains a single contiguous bitmap of size $N+1$, its auxiliary space requirement is \textbf{linear} $O(N)$. While acceptable for small to moderate values of $N$, this becomes prohibitive for very large limits.

\paragraph{Speedups:}
This implementation applies several well-known optimizations to improve efficiency:
\begin{itemize}[noitemsep]
    \item \textbf{Skip even numbers}: The algorithm treats 2 as a special case and processes only odd candidates (incrementing by 2), immediately halving the search space.
    \item \textbf{Start marking from $p^2$}: For each prime $p$, marking begins at $p^2$ rather than $2p$, since all smaller multiples have already been marked by smaller primes.
    \item \textbf{Mark only odd multiples}: When marking multiples of odd primes, the step size is $2p$ rather than $p$, skipping even composites.
\end{itemize}

After applying these optimizations, the total bit-work is estimated as:
\[
O\left(\frac{N}{2}\sum_{p=3}^{\sqrt N}\frac{1}{p}\right).
\]

\paragraph{Time Complexity:} Since the cost per prime for marking composites is proportional to $1/p$, and the sum $\sum_{p \le \sqrt{N}} 1/p$ grows as $O(\log \log N)$, the overall time complexity is $O(N \log \log N)$ with an improved constant factor (effectively $N/2$ instead of $N$ for the naive implementation).

\subsection{Segmented Sieve of Eratosthenes (SSoE)} \label{alg:ssoe}

The Segmented Sieve of Eratosthenes (SSoE) is a memory-efficient variant of the classical SoE algorithm. Rather than allocating a single monolithic bitmap of size $N$, SSoE divides the candidate range into smaller segments of size $\Delta$ (typically $\sqrt{N}$) and processes each segment sequentially. 

The algorithm begins by generating all \emph{root primes} up to $\sqrt{N}$ using the standard SoE. These root primes are then used to sieve each subsequent segment independently. This segmented approach preserves the time complexity of SoE while dramatically improving cache locality and reducing memory requirements, making it suitable for finding primes up to very large limits.

\begin{algorithm}[H]
    \caption{Segmented Sieve of Eratosthenes (SSoE)}
    \begin{algorithmic}[1]        
        \Procedure{SSoE}{$\ttt{int}: n$}{ $\rightarrow \ttt{INT\_ARRAY}$}
            \State \Comment{\textbf{1. Initialization:}}
            \State $segment\_size \gets \lceil\sqrt{n}\rceil$ \algocomment{compute segment size}
            \State \ttt{BITMAP:} $segment \gets \Call{bitmap\_init}{segment\_size + 1, 1}$ \algocomment{create segment bitmap}
            \State \ttt{INT\_ARRAY:} $primes \gets \Call{SoE}{segment\_size}$ \algocomment{initialize primes array with root primes up to $\sqrt{n}$}


            \State \Comment{\textbf{2. Process subsequent segments}}
            \State $low \gets segment\_size + 1$ \algocomment{initialize segment bounds}
            \State $high \gets low + segment\_size - 1$
            
            \While{$low \le n$} \algocomment{iterate over segments}
                \State \Call{bitmap\_set\_all}{$segment$} \algocomment{reset segment bitmap}
                
                \For{$p \ge 3 \in primes$} \algocomment{sieve current segment using root primes}
                    \If{$p^2 > high$} \textbf{break} \EndIf \algocomment{no need to process larger primes}
                    
                    \State $start \gets \max(p^2, \lceil low/p \rceil \times p)$ \algocomment{find first multiple of $p$ in current segment}
                    \State $start \gets start \textbf{ if } start \bmod 2 \neq 0 \textbf{ else } start + p$ \algocomment{ensure start is odd}
                    \State \Call{bitmap\_clear\_steps}{$segment, 2p, start - low, high - low$} \algocomment{mark multiples}
                \EndFor
                
                \For{$i \gets low$ to $high$ with step 2} \algocomment{collect primes from segment}
                    \If{\Call{bitmap\_get\_bit}{$segment, i - low$}} \algocomment{if unmarked, it's prime}
                        \State \Call{int\_array\_push}{$primes, i$}
                    \EndIf
                \EndFor
                
                \State $low \gets low + segment\_size$ \algocomment{move to next segment}
                \State $high \gets \min(high + segment\_size, n)$
            \EndWhile

            \State \Comment{\textbf{3. Return $primes$}}
            \State \Return \emph{primes}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\subsubsection*{Complexity Analysis}

\paragraph{Space Complexity:} SSoE requires only $O(\sqrt{N})$ auxiliary space for the segment bitmap, plus $O(\pi(\sqrt{N}))$ space for storing the root primes (where $\pi(x)$ denotes the prime counting function). This represents a significant improvement over SoE's $O(N)$ space requirement, enabling the algorithm to handle limits far beyond what would fit in main memory using a monolithic sieve.

\paragraph{Speedups:}
SSoE inherits all the optimizations from SoE: skipping even numbers, starting marking from $p^2$, and using step sizes of $2p$. The key additional advantage lies in improved cache locality—by processing the sieve in small, cache-resident segments, the algorithm achieves dramatically faster memory access patterns compared to the monolithic SoE bitmap. For each segment, we only use primes up to $\sqrt{high}$ for marking, where $high$ is the upper bound of the current segment.

\paragraph{Time Complexity:} The time complexity remains $O(N \log \log N)$, matching SoE asymptotically. However, in practice, SSoE often runs faster due to superior cache utilization, especially for large $N$ where the monolithic bitmap would thrash the cache.

\subsection{Sieve of Euler (SoEu)} \label{alg:soeu}

The Sieve of Euler, also known as the Linear Sieve, is an elegant prime-finding algorithm that ensures each composite number is marked exactly once. Unlike the Sieve of Eratosthenes, which may mark the same composite multiple times (for example, 15 is marked when sieving with both 3 and 5), the Sieve of Euler guarantees that each composite $n$ is marked only by its smallest prime factor (SPF).

This single-marking property makes the algorithm theoretically linear in the number of operations performed. However, in practice, it may be slower than optimized versions of SoE due to less predictable memory access patterns and the overhead of the inner loop over all collected primes.

\begin{algorithm}[H]
    \caption{Sieve of Euler (SoEu)}
    \begin{algorithmic}[1]        
        \Procedure{SoEu}{$\ttt{int}: n$}{ $\rightarrow \ttt{INT\_ARRAY}$}
            \State \Comment{\textbf{1. Initialization:}}
            \State \ttt{INT\_ARRAY:} $primes \gets \Call{int\_array\_init}$ \algocomment{initialize primes array}
            \State \Call{int\_array\_push}{$primes, 2$} \algocomment{add 2 to primes}
            \State \ttt{BITMAP:} $sieve \gets \Call{bitmap\_init}{n + 1, 1}$ \algocomment{create sieve bitmap}

            \State \Comment{\textbf{2. Sieve Process: collect primes and mark composites}}
            \For{$i \gets 3$ to $n$ with step 2} \algocomment{iterate over odd candidates}
                \If{\Call{bitmap\_get\_bit}{$sieve, i$}} \algocomment{if $i$ is prime}
                    \State \Call{int\_array\_push}{$primes, i$} \algocomment{add $i$ to primes}
                \EndIf
                
                \For{$p > 2 \in primes$} \algocomment{mark multiples using collected primes}
                    \If{$p \times i > n$} \textbf{break} \algocomment{stop if product exceeds $n$}
                    \EndIf 
                    \State \Call{bitmap\_clear\_bit}{$sieve, p \times i$} \algocomment{mark $p \times i$ as composite}
                    \If{$i \bmod p = 0$} \textbf{break}  \algocomment{crucial: stop at smallest prime factor}
                    \EndIf
                \EndFor
            \EndFor

            \State \Comment{\textbf{3. Return $primes$}}
            \State \Return \emph{primes}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\subsubsection*{Complexity Analysis}

\paragraph{Space Complexity:} The algorithm requires $O(N)$ space for the sieve bitmap and $O(\pi(N))$ space for storing the progressively collected primes list.

\paragraph{Speedups:}
The key optimization is the inner loop termination condition at line 13: when $i \bmod p = 0$, we know that $p$ is the smallest prime factor of $i$. Therefore, any composite of the form $i \times q$ for primes $q > p$ will have a smaller prime factor (namely $p$), and will be marked later when that smaller factor is encountered. This crucial insight ensures each composite is marked exactly once.

The algorithm skips even numbers by starting at 3 and incrementing by 2, processing only odd candidates. The total bit-work for marking all odd composites exactly once is:
\[
O\left(\frac{N}{2}\right).
\]

\paragraph{Time Complexity:} The algorithm achieves true $O(N)$ time complexity since each candidate number is visited at most once during the marking phase. However, the constant factors are typically higher than optimized SoE implementations due to the nested loop structure over all collected primes and less cache-friendly memory access patterns.

\subsection{Sieve of Sundaram (SoS)} \label{alg:sos}

The Sieve of Sundaram is an elegant algorithm for finding all prime numbers up to a specified limit $N$. Discovered by Indian mathematician S. P. Sundaram in 1934, it operates on a transformed space of indices that naturally correspond to odd numbers.

The algorithm works by eliminating numbers of the form $i + j + 2ij$ from consideration, where $1 \le i \le j$ and $i + j + 2ij \le k = \lfloor(N-1)/2\rfloor$. After this elimination process, the remaining indices $i$ in the range $[1, k]$ correspond to primes via the transformation $2i + 1$.

This approach is conceptually elegant and naturally produces only odd primes (with 2 handled as a special case). The algorithm operates on a reduced space of size approximately $N/2$, similar to the optimization in SoE that skips even numbers.

\begin{algorithm}[H]
    \caption{Sieve of Sundaram (SoS)}
    \begin{algorithmic}[1]        
        \Procedure{SoS}{$\ttt{int}: n$}{ $\rightarrow \ttt{INT\_ARRAY}$}
            \State \Comment{\textbf{1. Initialization:}}
            \State \ttt{INT\_ARRAY:} $primes \gets \Call{int\_array\_init}$ \algocomment{initialize primes array}
            \State \Call{int\_array\_push}{$primes, 2$} \algocomment{add 2 (special case)}

            \State $k \gets \lfloor\frac{n - 1}{2}\rfloor$ \algocomment{compute the size of the odd set up to $n$}
            \State \ttt{BITMAP:} $sieve \gets \Call{bitmap\_init}{k, 1}$ \algocomment{create sieve bitmap}
            
            \State \Comment{\textbf{2. Sieve Process: mark non-primes and collect primes}}
            \State $n\_sqrt \gets \lceil\sqrt{n}\rceil$ \algocomment{root limit}
            \For{$i \gets 1$ to $k$} \algocomment{iterate through Sundaram space}
                \If{\Call{bitmap\_get\_bit}{$sieve, i$}} \algocomment{if $i$ is unmarked}
                    \State $p \gets 2i + 1$ \algocomment{compute prime from index}
                    \State \Call{int\_array\_push}{$primes, p$} \algocomment{add prime}
                    
                    \If{$p < n\_sqrt$} \algocomment{if $p$ is root prime}
                        \State $x_p \gets p \times i + i$ \algocomment{first composite index: corresponds to $p^2$}
                        \State \Call{bitmap\_clear\_steps}{$sieve, p, x_p, k$} \algocomment{mark composites additively with step $p$}
                    \EndIf
                \EndIf
            \EndFor

            \State \Comment{\textbf{3. Return $primes$}}
            \State \Return \emph{primes}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\subsubsection*{Complexity Analysis}

\paragraph{Space Complexity:} The algorithm requires \textbf{linear} $O(N/2)$ space for the sieve bitmap, which is approximately half the size of the classical SoE bitmap.

\paragraph{Speedups:}
The classical Sieve of Sundaram uses a multiplicative approach to mark composites, targeting numbers of the form $i + j + 2ij$, which yields $O(N \log N)$ time complexity. However, this implementation leverages the \Xp-Identity (described in Section~\ref{alg:siz}) to transform the marking process into an additive operation with step size $p$.

This optimization, which recognizes that odd composites in the Sundaram space can be marked additively, reduces the bit-work to:
\[
O\left(\frac{N}{2}\sum_{p=3}^{\sqrt N}\frac{1}{p}\right),
\]
matching the performance of optimized SoE.

\paragraph{Time Complexity:} With the additive marking optimization, the time complexity becomes $O(N \log \log N)$ with constant factors similar to the basic SoE implementation.

\subsection{Sieve of Atkin (SoA)} \label{alg:soa}

The Sieve of Atkin is a modern, theoretically optimal algorithm for finding all prime numbers up to a specified limit $N$. Developed by A. O. L. Atkin and Daniel J. Bernstein in 2004, it employs a fundamentally different approach from the classical Eratosthenes sieve, based on deep results from algebraic number theory involving quadratic forms.

The algorithm proceeds in two main phases:
\begin{enumerate}[noitemsep]
    \item \textbf{Preprocessing}: Mark candidates by testing specific modular arithmetic conditions based on three quadratic forms ($4x^2 + y^2$, $3x^2 + y^2$, and $3x^2 - y^2$). Numbers satisfying these conditions an odd number of times are potentially prime.
    \item \textbf{Sieving}: Eliminate composites by clearing all odd multiples of squares of primes.
\end{enumerate}

The bit-flipping mechanism in the first phase accounts for numbers that satisfy multiple quadratic forms—a number is prime if and only if it satisfies the conditions an odd number of times.

\begin{algorithm}[H]
    \caption{Sieve of Atkin (SoA)}
    \begin{algorithmic}[1]        
        \Procedure{SoA}{$\ttt{int}: n$}{ $\rightarrow \ttt{INT\_ARRAY}$}
            \State \Comment{\textbf{1. Initialization:}}
            \State \ttt{INT\_ARRAY:} $primes \gets \Call{int\_array\_init}$ \algocomment{initialize primes array}
            \State \Call{int\_array\_push}{$primes, 2$} \algocomment{add 2 and 3 as special cases}
            \State \Call{int\_array\_push}{$primes, 3$}

            \State \ttt{BITMAP:} $sieve \gets \Call{bitmap\_init}{n + 1, 0}$ \algocomment{create sieve bitmap, initially all 0}

            \State \Comment{\textbf{2. Preprocessing: Mark potential primes using Atkin conditions}}            
            \For{$x \gets 1$ while $4x^2 < n$} \algocomment{Condition 1: $4x^2 + y^2 \equiv 1 \pmod{12}$ or $5 \pmod{12}$}
                \For{$y \gets 1$ while $4x^2 + y^2 \le n$} 
                    \State $b \gets 4x^2 + y^2$
                    \If{$b \bmod 12 = 1$ or $b \bmod 12 = 5$}
                        \State \Call{bitmap\_flip\_bit}{$sieve, b$} \algocomment{flip bit}
                    \EndIf
                \EndFor
            \EndFor
            
            \For{$x \gets 1$ while $3x^2 < n$} \algocomment{Condition 2: $3x^2 + y^2 \equiv 7 \pmod{12}$}
                \For{$y \gets 1$ while $3x^2 + y^2 \le n$}
                    \State $b \gets 3x^2 + y^2$
                    \If{$b \bmod 12 = 7$}
                        \State \Call{bitmap\_flip\_bit}{$sieve, b$}
                    \EndIf
                \EndFor
            \EndFor
            
            \For{$x \gets 1$ while $2x^2 < n$} \algocomment{Condition 3: $3x^2 - y^2 \equiv 11 \pmod{12}$ and $x > y$}
                \For{$y \gets x - 1$ down to $1$}
                    \State $b \gets 3x^2 - y^2$
                    \If{$b > n$} \textbf{break} \EndIf
                    \If{$b \bmod 12 = 11$}
                        \State \Call{bitmap\_flip\_bit}{$sieve, b$}
                    \EndIf
                \EndFor
            \EndFor

            \State \Comment{\textbf{3. Sieve: Eliminate composites by removing multiples of squares}}
            \State $n\_sqrt \gets \lceil\sqrt{n}\rceil$
            \For{$p \gets 5$ to $n\_sqrt$ with step 2} \algocomment{iterate over odd numbers $\ge 5$}
                \If{\Call{bitmap\_get\_bit}{$sieve, p$}} \algocomment{if $p$ survived, it's prime}
                    \State \Call{bitmap\_clear\_steps}{$sieve, 2p^2, p^2, n$} \algocomment{clear odd multiples of $p^2$}
                \EndIf
            \EndFor

            \State \Comment{\textbf{4. Collect remaining primes}}
            \For{$p \gets 5$ to $n$ with step 2}
                \If{\Call{bitmap\_get\_bit}{$sieve, p$}}
                    \State \Call{int\_array\_push}{$primes, p$}
                \EndIf
            \EndFor

            \State \Comment{\textbf{5. Return $primes$}}
            \State \Return \emph{primes}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\subsubsection*{Complexity Analysis}

\paragraph{Space Complexity:} The algorithm requires $O(N)$ space for the sieve bitmap.

\paragraph{Speedups:} 
The Sieve of Atkin's primary advantage stems from its use of quadratic forms to identify prime candidates more directly than the Eratosthenes approach. By testing specific modular conditions based on well-chosen quadratic forms, many numbers are classified as prime or composite without the repeated marking inherent to the Eratosthenes sieve.

The bit-flipping mechanism elegantly handles numbers that satisfy multiple quadratic forms. The final elimination phase is remarkably efficient, removing only odd multiples of squares of primes (rather than all multiples), resulting in a cost per prime of approximately $N/(2p^2)$.

The total bit-work can be approximated as:
\[
O\left(N\left(\frac{1}{4} + \frac{1}{6} + \frac{1}{12}\right) + \frac{N}{2}\sum_{p=5}^{\sqrt N}\frac{1}{p^2}\right) = O(N),
\]
where the first term represents the quadratic form tests and the second term (which converges) represents the square-multiple elimination.

\paragraph{Time Complexity:} The theoretical time complexity is \textbf{linear} $O(N)$, which is asymptotically superior to SoE's $O(N \log \log N)$. However, in practice for typical ranges (up to $10^{12}$), highly optimized implementations of SoE or its segmented variants often outperform SoA due to simpler operations, better cache utilization, and lower constant factors. The Sieve of Atkin's advantage becomes more pronounced for extremely large limits where the asymptotic improvement dominates.

\subsection{Sieve-iZ (SiZ)} \label{alg:siz}

The Sieve-iZ (SiZ) algorithm employs a wheel factorization optimization based on modulus 6. By recognizing that all primes greater than 3 must be of the form $6k \pm 1$, the algorithm restricts its search to the candidate space $\iZ = \{z \in \mathbb{N} : z \equiv \pm 1 \pmod{6}\}$, automatically excluding all multiples of 2 and 3.

This approach reduces space requirements from $N$ to approximately $N/3$ and lowers the cost per prime for marking composites to $N/(3p)$. The algorithm maintains two separate bitmaps—one for each residue class modulo 6—and employs the \Xp-Identity for efficient, additive composite marking.

We begin with the formal definitions and notations that underpin the SiZ family of algorithms:
\begin{definition}[The $\iZ$ set]    
    The set of all natural numbers coprime to 6, expressed as the union of two disjoint residue classes modulo 6:
    \[
    \iZ \;=\; \{ z\in\mathbb{N} : z\equiv \pm 1 \pmod{6} \} \;=\; \iZ^-\,\dot\cup\,\iZ^+,
    \]
    where
    \[
    \iZ^- = \{6x-1:x\in\mathbb{N}\} \quad\text{and}\quad \iZ^+ = \{6x+1:x\in\mathbb{N}\},
    \]
    with cardinality $|\iZ| = \frac{1}{3} |\mathbb{N}|$. Note that all primes $p > 3$ belong to $\iZ$.
\end{definition}

\noindent
To map between indices $x$ and values of the form $6x\pm1$, we define the $iZ$ function:

\begin{definition}[The $iZ$ function]
    \[
    \boxed{\ iZ(x \in\mathbb{N},\ i \in \{\pm 1\})=6x+i\ }.
    \]
    Here, $x$ is the \emph{index} within a residue class, and $i$ specifies which residue class ($i = -1$ for $\iZ^-$ and $i = +1$ for $\iZ^+$).
\end{definition}

\paragraph{X-Arrays.}
When sieving $\iZ$ up to a limit $N$, we maintain two separate bitmaps, one per residue class, each of length $x_n = \lceil N/6 \rceil$ bits:
\begin{itemize}[noitemsep]
    \item \ttt{X5[$x$]} encodes candidates in the $\iZ^-$ lane (numbers of the form $6x-1$),
    \item \ttt{X7[$x$]} encodes candidates in the $\iZ^+$ lane (numbers of the form $6x+1$).
\end{itemize}

\paragraph{The \Xp-Identity: Composite Marking within $\iZ$.} \label{sec:xp_id}

The \Xp-Identity provides a modular, index-based formula for marking all composites of a given prime $p$ within the $\iZ$ space using simple additive steps. Derived from Lemma \ref{lem:iZ} (Appendix~\ref{appendix:iz_lemma}).
\begin{definition}[\Xp-Identity]
    Let $p=iZ(x_p,i_p)\in\mathbb{P}$ be a prime. The indices of composites of $p$ in the X-Arrays are given by:
    \[
    \begin{aligned}
        \texttt{X5:}&\quad \big\{\ p\,x_q - i_p\,x_p\ :\ x_q\ge x_p \in\mathbb{N}\ \big\} \quad\text{(composites of $p$ in $\iZ^-$)},\\
        \texttt{X7:}&\quad \big\{\ p\,x_q + i_p\,x_p\ :\ x_q\ge x_p \in\mathbb{N}\ \big\} \quad\text{(composites of $p$ in $\iZ^+$)}.
    \end{aligned}
    \]
    The constraint $x_q \ge x_p$ ensures we start marking from $p \ge q$, avoiding duplicate marking of composites whose smaller cofactor $q$ would have already been processed.
\end{definition}

\begin{algorithm}[H]
    \caption{Sieve-iZ (SiZ)}
    \begin{algorithmic}[1]
        
        \Procedure{SiZ}{$\ttt{int}: n$}{ $\rightarrow \ttt{INT\_ARRAY}$}
            
            \State \Comment{\textbf{1. Initialization}}
            \State \ttt{INT\_ARRAY:} $primes \gets \Call{int\_array\_init}$ \algocomment{initialize primes array and add 2 and 3} 
            \State \Call{int\_array\_push}{$primes, 2$}
            \State \Call{int\_array\_push}{$primes, 3$}

            \State $x_n \gets \left\lfloor\tfrac{n+1}{6}\right\rfloor + 1$ \algocomment{iZ upper bound}
            \State \ttt{BITMAP:} $x5 \gets \Call{bitmap\_init}{x_n, 1}$ \algocomment{create bitmap $x5$ for representing $iZ-$ candidates}
            \State \ttt{BITMAP:} $x7 \gets \Call{bitmap\_init}{x_n, 1}$ \algocomment{create bitmap $x7$ for representing $iZ+$ candidates}
            
            \State \Comment{\textbf{2. Sieve: mark composites and collect primes}}
            \State $n\_sqrt \gets \lceil\sqrt{n}\rceil$ \algocomment{root limit}
            \For{$x \gets 1$ to $x_n$} \algocomment{iterate over iZ candidates}
                \If{\Call{bitmap\_get\_bit}{$x5, x$}} \algocomment{if $iZ-$ prime}
                    \State $p \gets iZ(x,-1)$ \algocomment{compute $p$}
                    \State \Call{int\_array\_push}{$primes, p$} \algocomment{append $p$ to primes}
                    \If{$p < n\_sqrt$} \algocomment{mark multiples of $p$ in both $x5$ and $x7$ using the $Xp$ identity}
                        \State \Call{bitmap\_clear\_steps}{$x5, p, p\cdot x + x, x_n$} 
                        \State \Call{bitmap\_clear\_steps}{$x7, p, p\cdot x - x, x_n$} 
                    \EndIf
                \EndIf
                \If{\Call{bitmap\_get\_bit}{$x7, x$}} \algocomment{if $iZ+$ prime, similar process with $i_p = 1$}
                    \State $p \gets iZ(x,1)$
                    \State \Call{int\_array\_push}{$primes, p$}
                    \If{$p < n\_sqrt$}
                        \State \Call{bitmap\_clear\_steps}{$x5, p, p\cdot x - x, x_n$}
                        \State \Call{bitmap\_clear\_steps}{$x7, p, p\cdot x + x, x_n$}
                    \EndIf
                \EndIf
            \EndFor

            \State \Comment{\textbf{3. Return} $primes$}
            \State \Return $primes$
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\subsubsection*{Complexity Analysis}

\paragraph{Space Complexity:} SiZ maintains two bitmaps, each of size $\lceil N/6 \rceil$ bits, for a total auxiliary space of \textbf{linear} $O(N/3)$. This represents a 3× improvement over the classical SoE.

\paragraph{Time Complexity:} Using additive marking within the $\iZ$ space via the \Xp-Identity, where candidates are of the form $6x \pm 1$, the total bit-work is:
\[
O\left(\frac{N}{3}\sum_{p=5}^{\sqrt N}\frac{1}{p}\right),
\]
yielding a time complexity of $O(N \log \log N)$ with significantly improved constant factors compared to classical SoE (effectively $N/3$ instead of $N/2$ for the optimized SoE).

\subsection{Segmented Sieve-iZm (SiZm)} \label{alg:sizm}

The Segmented Sieve-iZm (SiZm) algorithm represents the culmination of optimizations in this framework, extending the $\iZ$ basic-wheel structure ($6x\pm 1$) with another layer of extended wheel we call iZ-Matrix (iZm), resulting in a highly cache-efficient algorithm with minimal memory footprint and excellent constant factors.

\subsubsection*{Key Ideas, Definitions, and Notations}

Let $vx = VX_k$ be the product of the first $k$ primes greater than 3. The iZm structure is a two-dimensional representation of the $\iZ$ space with fixed horizontal width $vx$ and unbounded vertical height $vy$. Each cell in the iZm corresponds to a candidate number in $\iZ$, mapped as follows:
\begin{itemize}
    \item \textbf{iZm5} (representing $\iZ^-$): \quad a candidate $z$ at [$x, y$] maps to $iZ(vx\cdot y + x,\ -1)$.
    \item \textbf{iZm7} (representing $\iZ^+$): \quad a candidate $z$ at [$x, y$] maps to $iZ(vx\cdot y + x,\ +1)$.
\end{itemize}
In the SiZm algorithm, we construct the $iZm$ with maximum horizontal width $VX_6$ (the product of the first 6 primes greater than 3: 5, 7, 11, 13, 17, and 19). This choice balances cache efficiency and can be tuned based on the hardware architecture. 
This segmentation strategy provides several key advantages:
\begin{itemize}[noitemsep]
    \item \textbf{Cache efficiency}: Irrespective of $N$, segments fit comfortably in CPU cache, dramatically improving memory access patterns, and reducing latency compared to monolithic bitmaps.
    \item \textbf{Wheel factorization}: Pre-sieving small primes (5, 7, 11, 13, 17, 19) that divide $VX_6$ in a base template eliminates the most expensive marking operations.
    \item \textbf{Scalability}: The algorithm can handle arbitrarily large limits $N$ without requiring massive contiguous memory allocation.
\end{itemize}

To maintain clarity and modularity, we define several helper procedures that encapsulate reusable components of the algorithm:

\paragraph{Compute optimal segment size \emph{VX}.}

The \ttt{iZm\_compute\_limited\_vx} procedure calculates the maximum segment size $VX$ that does not exceed $\lfloor (n + 1) / 6 \rfloor$, computed as the product of the first $k$ primes greater than 3. This ensures the segment size is appropriate for the given limit.

\begin{algorithmic}[1]
    \State \ttt{INT\_ARRAY:} $s\_primes = [5, 7, 11, 13, 17, 19, 23, 29, 31, 37, 41, 43, 47]$ \algocomment{global list of primes $> 3$}
    \Procedure{iZm\_compute\_limited\_vx}{$n$, $k$}
        \State $x_n \gets \lfloor\frac{n + 1}{6}\rfloor$
        \State $vx \gets 1$
        \State $i \gets 0$
        \For{$p \in s\_primes$ while $i < k$} \algocomment{iterate over first $k$ primes greater than 3}
            \If{$vx \times p > x_n$} \textbf{break} \algocomment{stop if product exceeds $x_n$}
            \EndIf
            \State $vx \gets vx \times p$
            \State $i \gets i + 1$
        \EndFor
        \State \Return $vx$
    \EndProcedure
\end{algorithmic}

\paragraph{Construct pre-sieved base bitmaps.}

The \ttt{iZm\_construct\_vx\_base} procedure sieves the base bitmaps for a segment of size $VX$ using all primes that divide $VX$. These pre-sieved base bitmaps serve as templates for initializing each segment's active bitmaps, eliminating the need to mark these small primes during individual segment processing.

\begin{algorithmic}[1]
    \Procedure{iZm\_construct\_vx\_base}{$vx$, $base\_x5$, $base\_x7$}
        \For{$p \in s\_primes$ such that $p \mid vx$} \algocomment{iterate over primes dividing $vx$}
            \State $ip \gets 1 \textbf{ if } p \bmod 6 == 1 \textbf{ else } -1$
            \State $x_p \gets \lfloor\frac{p + 1}{6}\rfloor$
            \State $bitmap \gets base\_x5 \textbf{ if } ip == -1 \textbf{ else } base\_x7$ \algocomment{select bitmap containing $p$}
            \State \Call{bitmap\_clear\_bit}{$bitmap, x_p$} \algocomment{mark $p$ itself as composite within the template}
            \State \Comment{Mark composites of $p$ in base bitmaps using the \Xp-Identity}
            \State \Call{bitmap\_clear\_steps}{$base\_x5, p, p \cdot x_p - ip \cdot x_p, vx$} 
            \State \Call{bitmap\_clear\_steps}{$base\_x7, p, p \cdot x_p + ip \cdot x_p, vx$}
        \EndFor
    \EndProcedure
\end{algorithmic}

\paragraph{The \emph{Xp} solver for segmented marking.}

The \ttt{iZm\_solve\_for\_xp} procedure computes the first composite index of a prime $p$ within a given segment $y$, based on the \Xp-Identity. This solver is crucial for determining where to begin marking composites when processing each segment.

\begin{algorithmic}[1]
    \Procedure{iZm\_solve\_for\_xp}{$m\_id$, $p$, $vx$, $y$}
        \State $ip \gets 1 \textbf{ if } p \bmod 6 == 1 \textbf{ else } -1$
        \State $xp \gets \lfloor\frac{p + 1}{6}\rfloor$
        \If{$y == 0$}
            \State \Return $p \cdot xp + m\_id \cdot ip \cdot xp$  \algocomment{first composite index in segment 0}
        \EndIf
        \State $xp \gets xp \textbf{ if } m\_id == ip \textbf{ else } p - xp$ \algocomment{normalize $xp$ for cross-lane marking}
        \State $x \gets p - ((vx \cdot y - xp) \bmod p)$ \algocomment{solve for first hit in segment $y$}
        \State \Return $x$
    \EndProcedure
\end{algorithmic}

\begin{algorithm}[H]
    \caption{Segmented Sieve-iZm (SiZm)}
    \begin{algorithmic}[1]
        \Procedure{SiZm}{$\ttt{int}: n$}{ $\rightarrow \ttt{INT\_ARRAY}$}
            
            \State \Comment{\textbf{1. Initialization}}
            \State $vx \gets \Call{iZm\_compute\_limited\_vx}{n, 6}$ \algocomment{compute segment size, up to $VX_6$}
            \State \Comment{Initialize $primes$ with $SiZ(iZ(vx, 1))$ to sieve first segment and collect root primes}
            \State \ttt{INT\_ARRAY:} $primes \gets \Call{SiZ}{iZ(vx, 1)}$ 
            \State \Comment{Initialize base bitmaps, pre-sieved by primes dividing $vx$}
            \State \ttt{BITMAP:} $base\_x5 \gets \Call{bitmap\_init}{vx, 1}$ \algocomment{create base bitmap for $\iZ^-$}
            \State \ttt{BITMAP:} $base\_x7 \gets \Call{bitmap\_init}{vx, 1}$ \algocomment{create base bitmap for $\iZ^+$}
            \State $\Call{iZm\_construct\_vx\_base}{vx, base\_x5, base\_x7}$ \algocomment{pre-sieve base bitmaps}

            \State \Comment{\textbf{2. Process remaining segments for $y \in [1, y\_limit]$}}
            \State $x_n \gets \left\lfloor\tfrac{n+1}{6}\right\rfloor + 1$ \algocomment{$\iZ$ upper bound}
            \State $y\_limit \gets \lfloor x_n / vx \rfloor$ \algocomment{number of full segments}

            \For{$y \gets 1$ to $y\_limit$ (inclusive)}
                \State \Comment{{a. Initialize active bitmaps from pre-sieved base templates}}
                \State \ttt{BITMAP:} $x5 \gets \Call{bitmap\_clone}{base\_x5}$ \algocomment{clone for active $\iZ^-$ sieve}
                \State \ttt{BITMAP:} $x7 \gets \Call{bitmap\_clone}{base\_x7}$ \algocomment{clone for active $\iZ^+$ sieve}
        
                \State \Comment{{b. Mark composites using root primes in current segment}}
                \State $x\_limit \gets vx \textbf{ if } (y < y\_limit) \textbf{ else } x_n \bmod vx$ \algocomment{handle partial last segment}
                \State $yvx \gets y \times vx$ \algocomment{base index offset for current segment}
                \State $root\_limit \gets \sqrt{6 \cdot (yvx + x\_limit)} + 1$ \algocomment{local root limit}
                \For{$p \ge 5 \in primes$} \algocomment{iterate over primes $> 3$}
                    \If{$p \mid vx$} \textbf{continue} \algocomment{skip pre-sieved primes}
                    \EndIf
                    \If{$p > root\_limit$} \textbf{break} \algocomment{no need for larger primes}
                    \EndIf
                    \State \Comment{Mark composites of $p$ in current segment using \Xp~solver}
                    \State \Call{bitmap\_clear\_steps}{$x5, p, iZm\_solve\_for\_xp(-1, p, vx, y), x\_limit$} 
                    \State \Call{bitmap\_clear\_steps}{$x7, p, iZm\_solve\_for\_xp(1, p, vx, y), x\_limit$}
                \EndFor

                \State \Comment{{c. Collect primes from current segment}}   
                \For{$x \gets 1$ to $x\_limit$ (inclusive)}
                    \If{\Call{bitmap\_get\_bit}{$x5, x$}} \algocomment{unmarked index in $\iZ^-$}
                        \State \Call{int\_array\_push}{$primes, iZ(yvx + x, -1)$}
                    \EndIf

                    \If{\Call{bitmap\_get\_bit}{$x7, x$}} \algocomment{unmarked index in $\iZ^+$}
                        \State \Call{int\_array\_push}{$primes, iZ(yvx + x, 1)$}
                    \EndIf
                \EndFor
            \EndFor

            \State \Comment{\textbf{3. Return} $primes$}
            \State \Return $primes$
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\subsubsection*{Complexity Analysis}

\paragraph{Space Complexity:} 
Rather than using an arbitrary segment size of $\sqrt{N}$ as in traditional segmented sieves, SiZm decouples the segment size from $N$ and employs a fixed maximum wheel size:
$$VX_6 = 5 \times 7 \times 11 \times 13 \times 17 \times 19 = 1{,}616{,}615.$$

Beyond the $O(\pi(\sqrt{N}))$ space for storing root primes, each segment requires two bitmaps of maximum size $VX_6$ (approximately 1.6 million bits $\approx$ 200 KB each), yielding a space complexity of $O(\pi(\sqrt{N})) + O(1)$. Crucially, this space requirement is \emph{independent of $N$}, ensuring that segments remain cache-resident even for extremely large limits.

\paragraph{Speedups:} 
Building upon the SiZ optimizations, SiZm achieves further improvements by pre-sieving the small primes that divide $VX_6$ (namely 5, 7, 11, 13, 17, and 19) in the wheel structure. Since these small primes contribute disproportionately to the total marking cost (due to their high frequency), eliminating them from the per-segment marking phase yields substantial performance gains.

After pre-sieving these primes, the remaining bit-work is estimated as:
\[
O\left(\frac{N}{3}\sum_{p=23}^{\sqrt N}\frac{1}{p}\right) + O(1),
\]
where the $O(1)$ term represents the one-time cost of wheel construction, which is negligible compared to the total work.

\paragraph{Time Complexity:} The time complexity remains $O(N \log \log N)$, but with dramatically improved constant factors. In practice, SiZm often outperforms all other algorithms in this document for large limits, due to the combination of reduced search space, pre-sieving of expensive small primes, and excellent cache locality.



\section{Summary}

Table~\ref{tab:prime_sieves_summary} provides a comprehensive comparison of the prime sieve algorithms presented in this document. The algorithms exhibit a range of trade-offs between space efficiency, time complexity, and practical performance.

\textbf{Key observations:}
\begin{itemize}[noitemsep]
    \item \textbf{Space vs. Time}: While SoEu and SoA achieve linear $O(N)$ time complexity, they require $O(N)$ space. In contrast, SSoE and SiZm achieve similar practical performance with dramatically reduced space requirements.
    \item \textbf{Segmentation benefits}: SSoE and SiZm demonstrate that segmentation provides substantial practical advantages through improved cache locality, even when asymptotic complexity remains unchanged.
    \item \textbf{Wheel factorization}: The progression from SoE ($O(N)$ space) to SiZ ($O(N/3)$ space) to SiZm ($O(1)$ space beyond root primes) shows the power of wheel factorization combined with segmentation.
    \item \textbf{Practical considerations}: Despite superior asymptotic complexity, SoEu and SoA often underperform optimized implementations of SoE-family algorithms in practice due to cache effects and instruction-level overhead.
\end{itemize}

\begin{table}[H]
    \centering
    \renewcommand{\arraystretch}{2}
    \begin{tabular}{l|c|c|c}
        \hline
        \textbf{Algorithm} & \textbf{Space Complexity} & \textbf{Time Complexity} & \textbf{Bit-Work} \\
        \hline
        SoE & $O(N)$ & $O(N \log \log N)$ & $O\left(\frac{N}{2}\sum_{p=3}^{\sqrt N}\frac{1}{p}\right)$ \\
        SSoE & $O(\pi(\sqrt{N}) + \sqrt{N})$ & $O(N \log \log N)$ & $O\left(\frac{N}{2}\sum_{p=3}^{\sqrt N}\frac{1}{p}\right)$ \\
        SoEu & $O(N)$ & $O(N)$ & $O\left(\frac{N}{2}\right)$ \\
        SoS & $O(N/2)$ & $O(N \log \log N)$ & $O\left(\frac{N}{2}\sum_{p=3}^{\sqrt N}\frac{1}{p}\right)$ \\
        SoA & $O(N)$ & $O(N)$ & $O\left(\frac{N}{2} + \frac{N}{2}\sum_{p=5}^{\sqrt N}\frac{1}{p^2}\right)$ \\
        SiZ & $O(N/3)$ & $O(N \log \log N)$ & $O\left(\frac{N}{3}\sum_{p=5}^{\sqrt N}\frac{1}{p}\right)$ \\
        SiZm & $O(\pi(\sqrt{N})) + O(1)$ & $O(N \log \log N)$ & $O\left(\frac{N}{3}\sum_{p=23}^{\sqrt N}\frac{1}{p}\right) + O(1)$ \\
        \hline
    \end{tabular}
    \caption{Comprehensive comparison of prime sieve algorithms. Space complexity represents auxiliary space beyond the output. Bit-work estimates reflect the number of bit operations performed during sieving.}
    \label{tab:prime_sieves_summary}
\end{table}

\newpage
\appendix

\section{Algebraic Study of the iZ Function} \label{appendix:iz_lemma}

This appendix provides the mathematical foundation for the \Xp-Identity used in the SiZ and SiZm algorithms. We establish a general lemma about products in the $\iZ$ space and derive the key equivalence relation that enables efficient composite marking.

\medskip
\begin{lemma}
    \label{lem:iZ}
    Let $m = 6$ be the modulus defining the $\iZ$ space.
    Let $\mathcal{I} = \{\pm 1\}$ be the set of residues coprime to $m$, which is closed under multiplication modulo $m$.
    Let $iZ(x \in \mathbb{N}, i \in \mathcal{I}) = mx + i$.

    \noindent 
    For any $p=iZ(x_p,i_p)$ and $q=iZ(x_q,i_q)$ in $\iZ$, their product $z=pq$ can be expressed as $z=iZ(x_z,i_z)$, where
    \[
    x_z = m x_p x_q + i_p x_q + i_q x_p
    \quad\text{and}\quad
    i_z = i_p i_q \in \mathcal{I}.
    \]

    \noindent 
    Consequently, for each prime factor $p$ that divides $z$, the following equivalence holds:
    \[
    p \mid z \quad \Longleftrightarrow \quad x_z \equiv \left(\frac{i_z}{i_p}\right) x_p \pmod{p}.
    \]
    
    \noindent
    This equivalence forms the basis of the \Xp-Identity, enabling efficient additive marking of composites in the $\iZ$ space.
\end{lemma}

\begin{proof} \label{proof:iZ}
We expand the product $z = pq$ by substituting the $iZ$ representations:
\[
z = pq = (m x_p + i_p)(m x_q + i_q) = m^2 x_p x_q + m i_p x_q + m i_q x_p + i_p i_q.
\]

\noindent 
Factoring out $m$ from the first three terms:
\[
z = m (m x_p x_q + i_p x_q + i_q x_p) + i_p i_q = m x_z + i_z,
\]
which confirms that $z \in \iZ$ with $z = iZ(x_z, i_z)$, where:

\begin{itemize}[leftmargin=*,noitemsep]
    \item The \emph{product residue} $i_z \in \mathcal{I}$ is determined by:
    $$
    i_z = i_p i_q = \begin{cases}
    +1 & \text{if } i_p = i_q,\\
    -1 & \text{if } i_p \neq i_q.
    \end{cases}
    $$
    
    \item The \emph{product index} $x_z$ can be rewritten in terms of the original factors:
    \begin{align*}
    x_z &= m x_p x_q + i_p x_q + i_q x_p \\
        &= (m x_p + i_p) x_q + i_q x_p \\
        &= p x_q + i_q x_p \\
        &= (m x_q + i_q) x_p + i_p x_q \\
        &= q x_p + i_p x_q.
    \end{align*}
\end{itemize}

\noindent
From the first form $x_z = p x_q + i_q x_p$, we obtain:
$$
x_z \equiv i_q x_p \pmod{p}.
$$

\noindent
Multiplying both sides by $i_p$ (noting that $i_p \cdot i_q = i_z$ and $i_p^2 = 1$ since $i_p \in \{\pm 1\}$):
$$
i_p x_z \equiv i_p i_q x_p \equiv i_z x_p \pmod{p}.
$$

\noindent
Since $\gcd(i_p, p) = 1$ (as $i_p \in \{\pm 1\}$), we can equivalently write:
$$
x_z \equiv \frac{i_z}{i_p} x_p \pmod{p},
$$
which establishes the desired equivalence relation.

\medskip
\noindent
\textbf{Application to the \Xp-Identity:} This equivalence allows us to identify all composite indices in the X-arrays by solving for $x_z$ modulo $p$. For a prime $p = iZ(x_p, i_p)$, its multiples in each residue class can be marked additively with step size $p$, starting from the appropriate initial index determined by this relation. This forms the mathematical foundation for the efficient composite marking in SiZ and SiZm algorithms.
\end{proof}



\begin{thebibliography}{9}
    \bibitem{SoE}
    Wikipedia contributors. Sieve of Eratosthenes. In \textit{Wikipedia, The Free Encyclopedia}. Retrieved January 1, 2026, from \url{https://en.wikipedia.org/wiki/Sieve_of_Eratosthenes}

    \bibitem{SSoE}
    Wikipedia contributors. Segmented sieve. In \textit{Wikipedia, The Free Encyclopedia}. Retrieved January 1, 2026, from \url{https://en.wikipedia.org/wiki/Sieve_of_Eratosthenes#Segmented_sieve}

    \bibitem{SoEu}
    Wikipedia contributors. Euler's sieve. In \textit{Wikipedia, The Free Encyclopedia}. Retrieved January 1, 2026, from \url{https://en.wikipedia.org/wiki/Sieve_of_Eratosthenes#Euler's_sieve}

    \bibitem{SoS}
    Wikipedia contributors. Sieve of Sundaram. In \textit{Wikipedia, The Free Encyclopedia}. Retrieved January 1, 2026, from \url{https://en.wikipedia.org/wiki/Sieve_of_Sundaram}

    \bibitem{SoA}
    Wikipedia contributors. Sieve of Atkin. In \textit{Wikipedia, The Free Encyclopedia}. Retrieved January 1, 2026, from \url{https://en.wikipedia.org/wiki/Sieve_of_Atkin}
\end{thebibliography}


\end{document}
